---
title: "Introduction to Quarto (seminar 1)"
author: "SPACIAL Unit"
format: html
editor: visual
---

### Today's program:

1.  Load libraries

2.  Set working directory

3.  Import data

4.  Explore data

    a)  Explore data types in the data set
    b)  Which variables can be found in the data set?

5.  Questions on the data

    a)  How many unique plots in the data?
    b)  Which counties are in the data?
    c)  How many plots per county?
    d)  Find if there's NAs in county_region column?
        i)  Change the NA values
    e)  What data type is the "plant_date" column?
        i)  Which unique planting years are in the data?
        ii) How many planted in each year per county?

6.  Data exploration through visualization

    a)  Visualize the number of plots per county?
        i)  Change the color and width of the the bars
        ii) Look up some information on the type of plot you used
    b)  Show the distribution in size of the plots in different counties
    c)  Plot the number trees planted vs the number survived
        i)  Add an extra variable as aesthetic

7.  Data manipulation (for next seminar)

## 1. Libraries

```{r}
# Load libraries

```

## 2. Set working directory

## 3. Import data

```{r}
# Import data

```

## 4. Explore data

```{r}
# Explore data types in the data set

# Which variables can be found in the data set
```

## 5. Answer the questions

```{r}
# How many unique plots in the data?

```

```{r}
# Which counties are in the data?

```

```{r}
# How many plots per county?

```

```{r}
# Find NAs in the county_region column and change the value

```

```{r}
# What data type is the "plant_date" column?
# Which unique planting years are in the data?
# How many planted in each year per county

```

## 6. Data exploration by visualization

```{r}
# Visualize the number of plots per county
  # Change the color and width of the the bars
  # Look up some information on the type of plot you used (eg what does 'identity' mean?)

```

```{r}
# Show the distribution in size of the plots in different counties
  # Does the y limit make sense?

```

```{r}
# Plot the number trees planted vs the number survived
  #Add an extra variable as aesthetic 

```

## 7. Data manipulation (next seminar)

```{r}
# Create new column showing the survival in % and plot


```
