---
title: "Advanced data manipulation (Seminar 3)"
author: "SPACIAL Unit"
format: html
editor: visual
---

### Today's (30/11/2022) program:

-  Recap last seminar

-  Publish html with quarto

-  Continue on data manipulation (Seminar 3 - 30/11/2022)

    7. Data transformations
    8. Specific transformations
    9. Apply data manupilation knowledge
    10. Base R functions such as loops (optional)

## 1. Libraries

```{r}
# Load libraries
library(tidyverse)
library(data.table)

# If not installed yet:
install.packages('tidyverse')
install.packages('data.table')
```

## 2. Set working directory

## 3. Import data

```{r}
# Import data
ra <- read_csv('Rwanda_subset.csv')

```

## 4. Explore data

```{r}
# Explore data types in the data set
summary(ra)
View(ra)
names(ra)
# Which variables can be found in the data set
```

## 5. Answer the questions

```{r}
# How many unique plots in the data?
length(unique(ra$plotID))
```

```{r}
# Which counties are in the data?
unique(ra$county)
```

```{r}
# How many plots per county?
ra %>% 
  group_by(county) %>% 
  summarise(number_plots=n())
  
```

```{r}
# Find NAs in the county_region column and change the value
length(which(is.na(ra$county_region)==T))

```

```{r}
# What data type is the "plant_date" column?
class(ra$plant_date)

# Which unique planting years are in the data?
unique(year(ra$plant_date))
# How many planted in each year per county
ra %>% 
  group_by(county, year(plant_date)) %>% 
  summarise(number_planted=n())

```

## 6. Data exploration by visualization

```{r}
# Visualize the number of plots per county
ra %>% 
  group_by(county) %>% 
  summarise(number_planted=n()) %>% 
  ggplot()+
  geom_col(aes(x=county, y=number_planted), fill='darkblue')+
  labs(y='Number planted', x='County')
  # Change the color and width of the the bars
  # Look up some information on the type of plot you used (eg what does 'identity' mean?)

```

```{r}
# Show the distribution in size of the plots in different counties
ggplot(ra)+
  geom_histogram(aes(x=Hectare))+
  facet_wrap(~county)

ggplot(ra)+
  geom_boxplot(aes(x=county, y=Hectare))+
  ylim(0,1)

ggplot(ra)+
  geom_violin(aes(x=county, y=Hectare), fill='grey',
              col='dark green')+
  ylim(0,1)

# Does the y limit make sense?

```

```{r}
# Plot the number trees planted vs the number survived
ggplot(ra)+
  geom_point(aes(x=number_planted, y=number_survived, col=Hectare))+
  geom_smooth(aes(x=number_planted,y=number_survived),
              method='loess')+
  ylim(0,200)+
  xlim(0,200)

ggplot(ra)+
  geom_point(aes(x=number_planted, y=number_survived, col=Hectare))+
  geom_smooth(aes(x=number_planted,y=number_survived),
              method='lm')+
  ylim(0,200)+
  xlim(0,200)

  #Add an extra variable as aesthetic 

```

# Conintue on data manipulation (Seminar 3 - 30/11/2022)

## 7. Data transformations

```{r}
# Create new column and add to the dataset



# Show the survival ratio and plot



# Show survival per county



# Create a new variable with the number of trees per hectare and visualize against survive_ratio



```
## 8. Specific transformations

```{r}
# Create new column that equals 1 when "Hectare" is >0.5ha and "number_planted" is <10



# See in how many instances this is the case



# Explore the if_else function


# Find NAs values in "county_region"


# Use the if_else function to change the NA into "No region"



# Remove the rows with "No region" in "country_region"




# Try to combine as many as the commands above in one pipe



```

## 9. Apply data manipulation knowledge
```{r}
# Find the plots with the highest 10% number of trees per hectare in a garden. Then, exclude all columns but plotID, county, plant_date and trees_hectares. Then visualise the result 



```

## 10. Introduction to functions and loops (optional)

```{r}
# Create new column that equals 1 when "Hectare" is >0.5ha and "number_planted" is <10




# See in how many instances this is the case



```